From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-04-21 paper.tex
Draft messaging section

\begin{abstract}

  The Blue Brain project has been investing in high-performance visualization
  (HPV) to complement its HPC strategy since its inception in 2007. In 2011,
  this strategy has been accelerated to develop innovative visualization
  solutions through through increased funding and strategic partnerships with
  other research institutions.

  We present the key elements of this HPV ecosystem, which integrates C++
  visualization applications with novel collaborative display systems. We
  motivate how our strategy of transforming high-performance visualization
  engines into services enables a variety of use cases, not only for the
  integration with high-fidelity displays, but also to build service oriented
  architectures, to link into web applications and to provide remote services to
  python applications.

\end{abstract}

\section{Collaborative Display Systems}

@@

\begin{figure}[h!t]
  \includegraphics[width=\textwidth]{images/tdw}
  \caption{\label{fTDW}Blue Brain $4\times 3$ tiled display wall}
\end{figure}

@@

\begin{figure}[h!t]
  \includegraphics[width=\textwidth]{images/opendeck}
  \caption{\label{fOpenDeck}OpenDeck concept rendering}
\end{figure}

@@

\begin{figure}[h!t]
  \includegraphics[width=\textwidth]{images/tideweb}
  \caption{\label{fTideWeb}Tide web interface}
\end{figure}

@@

\section{Messaging/Services}

>>>>

\begin{abstract}
  The Blue Brain project has been investing in high-performance visualization
  (HPV) to complement its HPC strategy since its inception in 2007. In 2011,
  this strategy has been accelerated to develop innovative visualization
  solutions through through increased funding and strategic partnerships with
  other research institutions.

  We present the key elements of this HPV ecosystem, which integrates C++
  visualization applications with novel collaborative display systems. We
  motivate how our strategy of transforming high-performance visualization
  engines into services enables a variety of use cases, not only for the
  integration with high-fidelity displays, but also to build service oriented
  architectures, to link into web applications and to provide remote services to
  python applications.
\end{abstract}

\section{Collaborative Displays}

@@

\begin{figure}[h!t]
  \includegraphics[width=\columnwidth]{images/tdw}
  \caption{\label{fTDW}Blue Brain $4\times 3$ tiled display wall}
\end{figure}

@@

\begin{figure}[h!t]
  \includegraphics[width=\columnwidth]{images/opendeck}
  \caption{\label{fOpenDeck}OpenDeck concept rendering}
\end{figure}

@@

\begin{figure}[h!t]
  \includegraphics[width=\columnwidth]{images/tideweb}
  \caption{\label{fTideWeb}Tide web interface}
\end{figure}

@@

\section{C++ Messaging and Services}

All Blue Brain applications integrate a messaging framework which allows them
to be used as services in a variety of use cases. For example, the Tide web
server providing the user interface shown in \fig{fTideWeb}, is based on this
messaging solution. Other use cases are remote python APIs, JavaScript user
interfaces and service architectures combining multiple visualization
applications with data providers such as HPC simulations.

The base communication layer ZeroEQ utilizes ZeroMQ as the transport layer, the
ZeroConf protocol for discovery, and our novel ZeroBuf serialization library for
high-performance messaging. A fully integrated HTTP server provides a bridge to
JavaScript, Python and similar environments by implementing REST APIs with JSON
payload. \fig{fUML} shows a class diagram of our messaging solution.

\begin{figure*}[ht]\center
  \includegraphics[width=\columnwidth]{images/ZeroMS}
  \caption{\label{fUML}UML Diagram of the main messaging classes}
\end{figure*}

\subsection{ZeroEQ}

ZeroEQ is the base messaging library, wrapping up existing technologies into an
API which is convenient to use and easy to integrate into C++ code. It provides
two messaging services: publish-subscribe and HTTP. For binary and JSON
encodings it relies on a simple \textsf{Serializable} interface, for which
ZeroBuf provides a sample implementation. To facilitate the simple use case of
linking a few applications, ZeroEQ uses the zeroconf protocol to discover and
connect to related applications. For more complex scenarios, explicit
connections are supported.

\subsubsection{Publish-Subscribe}

The publish-subscribe service is implemented in a \textsf{Publisher} and
\textsf{Subscriber} class. It provides event-based messaging, based on a 128-bit
message type with arbitrary payload. The message type is used for message
subscription, filtering and routing. The payload is expected to be uniquely
identified by the message type, that is, all applications agree for the decoding
and semantics of any given message type. The ZeroBuf section outlines how this
is implemented. The underlying transport uses ZeroMQ pub-sub sockets.

The pub-sub service is stateless, that is, applications have no expectation of
when messages are received or who receives published messages. This
communication pattern naturally leads to robust services, since there is no
possiblity for deadlocks or undefined behaviour. The pub-sub API is provided in
two flavors: a simple $pointer \& size$ memory buffer, and a higher level
object-based abstraction. The object-based API is syntactic sugar for the
low-level API, and allows automatic publish and update of objects with a few
lines of code. It uses the \textsf{toBinary()} and \textsf{fromBinary()} methods
of the \textsf{Serializable} interface to call the low-level API.

\begin{lstlisting}[float, caption=Publish-Subscribe Example, label=lPubSub]
zerobuf::render::Camera camera;
zeroeq::Publisher publisher;
zeroeq::Subscriber subscriber;

subscriber.subscribe( camera );

while( rendering )
{
    subscriber.receive( 0 /*ms, poll*/ );
    updateCamera( camera );
    publisher.publish( camera );
    renderFrame( camera );
}
\end{lstlisting}

The example in \lref{lPubSub} shows the integration of camera
synchronization in a visualization application. This example relies on the
builtin zeroconf protocol to connect application instances. Subscribers only
subscribe to events from publishers within the same session. The default
session name is the user name, and can be customized using an environment
variable or non-default constructor. Similarly, the subscriber can subscribe by
session or address. \lref{lPubSubExp} illustrates an explicitly addressed
subscription. Notice that the subscriber uses the publisher URI, which will
contain the concrete port chosen for the publisher.

\begin{lstlisting}[float, caption=Explicit Addressing,label=lPubSubExp]
zeroeq::URI uri( "tcp://localhost" );
zeroeq::Publisher publisher( uri, zeroeq::NULL_SESSION ); // deactivate zeroconf
zeroeq::Subscriber subscriber( publisher.getURI( )); // use concrete port
\end{lstlisting}

Subscribers are derived from a \textsf(Receiver) base class, which is shared
with the http server. All receivers can share their \textsf{receive()}
operation at construction time, that is, the blocking receive operation applies
to all receivers in the shared group. \lref{lSubShare} shows an example of
selectively receiving different updates on different input sockets.

\begin{lstlisting}[float,caption=Subscriber Sharing, label=lSubShare]
zeroeq::Subscriber local( zeroeq::URI( "localhost:29387" ));
zeroeq::Subscriber global( local );

local.subscribe( colorMap );
global.subscribe( camera );

while( true )
    local.receive(); // updates colorMap and camera
\end{lstlisting}

\subsubsection{HTTP Server}

The http server is built using cppnetlib \cite{cppnetlib} for the transport and
http protocol handling. It supports all standard http verbs (GET, POST, PUT,
PATCH, DELETE). It is a \textsf{zeroeq::Receiver}, that is, it can share its
\textsf{receive()} update operation with other subscribers and http servers.
Unlike a subscriber, the http server follows the HTTP request-reply semantics,
that is, a request received by a server has to be followed directly by its
reply. future

vocabulary
